---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  fig.retina = 2
)

library(patchwork)
library(tibble)
library(magrittr)
library(ggplot2)
library(gghilbertstrings)
```

# gghilbertstrings

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/Sumidu/gghilbertstrings.svg?branch=master)](https://travis-ci.com/Sumidu/gghilbertstrings)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Sumidu/gghilbertstrings?branch=master&svg=true)](https://ci.appveyor.com/project/Sumidu/gghilbertstrings)
[![Codecov test coverage](https://codecov.io/gh/Sumidu/gghilbertstrings/branch/master/graph/badge.svg)](https://codecov.io/gh/Sumidu/gghilbertstrings?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of gghilbertstrings is to ...

## Installation

You can install the released version of gghilbertstrings from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("gghilbertstrings")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Sumidu/gghilbertstrings")
```
## Usage

The `gghilbertstrings` package comes with functions for fast plotting of Hilbert curves in ggplot. At it's core is a fast RCPP implementation that maps a 1D vector to a 2D position.

The `gghilbertplot` function creates a Hilbert curve and plots individual data points to the corners of this plot. It automatically rescales the used `ID`-variable to the full range of the Hilbert curve. The method also automatically picks a suitable level of detail able to represent all values of `ID`.

The following figure shows different hilbert curves for different maximum `ID`s.
```{r hilbert, echo=FALSE}
p1 <- tibble(id = c(1,4)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 4") +  theme_void()
p2 <- tibble(id = c(1,16)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 16")  + theme_void()
p3 <- tibble(id = c(1,64)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 64")  + theme_void()
p4 <- tibble(id = c(1,256)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 256")  + theme_void()

p1 + p2 + p3 + p4 + plot_annotation(title = "Different depths of Hilbert curves")
```


### Plotting random data

The most simple way to plot data is to generate an `id` column that ranges from 1 to n, where n is the largest value to use in the Hilbert curve. Beware: The `id`s are rounded to integers.

```{r example}
library(gghilbertstrings)

# val is the ID column used here
df <- tibble(val = 1:128, 
       size = runif(128, 1, 5),        # create random sizes 
       color = rep(c(1,2,3,4),32))     # create random colours

gghilbertplot(df, val, 
                color = factor(color), # render color as a factor
                size = size, 
                add_curve = T)         # also render the curves
```


### Performance
```r 
# Performance benchmark
library(microbenchmark)
library(HilbertCurve)
library(gghilbertstrings)

# Compare the creation of coordinate systems
mb <- microbenchmark(times = 100, 
HilbertCurve = {
  hc <- HilbertCurve(1, 128, level = 7)
},
gghilbertstrings = {
  ggh <- hilbertd2xy(n = 2^7, df$val)
})
```



```{r eval=TRUE, include=FALSE}

# Performance benchmark
library(microbenchmark)
library(HilbertCurve)
library(tidyverse)
library(gghilbertstrings)
benchmark_times <- 100

# Compare the creation of coordinate systems
mb <- microbenchmark::microbenchmark(times = benchmark_times, 
  HilbertCurve = {
    hc = HilbertCurve(1, 128, level = 7)
},
  gghilbertstrings = {
    ggh <- hilbertd2xy(n = 2^7, df$val )
  })

```

```{r output, echo=FALSE, message=FALSE, warning=FALSE}
autoplot(mb) + 
  coord_flip() + 
  ggtitle("Comparison of runtime performance using 100 repetions on 7 levels") +
  labs(caption = "X-Axis on log-scale")
  
```


### Useful example

Link: https://www.kaggle.com/eliasdabbas/search-engine-results-flights-tickets-keywords under License CC0

