---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  fig.retina = 2
)

library(tidyverse)
library(patchwork)
library(tibble)
library(magrittr)
library(ggplot2)
library(gghilbertstrings)
library(glue)

reps <- 10 #takes about 50 seconds on iMac i9 3.5GHz
size_exponent <- 11
```

# gghilbertstrings

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/Sumidu/gghilbertstrings.svg?branch=master)](https://travis-ci.com/Sumidu/gghilbertstrings)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Sumidu/gghilbertstrings?branch=master&svg=true)](https://ci.appveyor.com/project/Sumidu/gghilbertstrings)
[![Codecov test coverage](https://codecov.io/gh/Sumidu/gghilbertstrings/branch/master/graph/badge.svg)](https://codecov.io/gh/Sumidu/gghilbertstrings?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

A [Hilbert curve](https://en.wikipedia.org/wiki/Hilbert_curve) (also known as a Hilbert space-filling curve) is a continuous fractal space-filling curve first described by the German mathematician David Hilbert in 1891, as a variant of the space-filling Peano curves discovered by Giuseppe Peano in 1890 (from Wikipedia).

This package provides an easy access to using Hilbert curves in `ggplot2`. 

## Installation

<!--
You can install the released version of gghilbertstrings from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("gghilbertstrings")
```
-->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Sumidu/gghilbertstrings")
```
## Usage

The `gghilbertstrings` package comes with functions for fast plotting of Hilbert curves in ggplot. At it's core is a fast RCpp implementation that maps a 1D vector to a 2D position.

The `gghilbertplot` function creates a Hilbert curve and plots individual data points to the corners of this plot. It automatically rescales the used `ID`-variable to the full range of the Hilbert curve. The method also automatically picks a suitable level of detail able to represent all values of `ID`.

The following figure shows different hilbert curves for different maximum `ID`s.
```{r hilbert, echo=FALSE}
p1 <- tibble(id = c(1,4)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 4") +  theme_void()
p2 <- tibble(id = c(1,16)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 16")  + theme_void()
p3 <- tibble(id = c(1,64)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 64")  + theme_void()
p4 <- tibble(id = c(1,256)) %>% gghilbertplot(id, add_curve = T) + ggtitle("n = 256")  + theme_void()

p1 + p2 + p3 + p4 + plot_annotation(title = "Different depths of Hilbert curves")
```


### Plotting random data

The most simple way to plot data is to generate an `id` column that ranges from 1 to n, where n is the largest value to use in the Hilbert curve. Beware: The `id`s are rounded to integers.

```{r example}
library(gghilbertstrings)

# val is the ID column used here
df <- tibble(val = 1:256, 
       size = runif(256, 1, 5),        # create random sizes 
       color = rep(c(1,2,3,4),64))     # create random colours

gghilbertplot(df, val, 
                color = factor(color), # render color as a factor
                size = size, 
                add_curve = T)         # also render the curves
```


### Performance
We run the creation of a coordinate system `r reps` times. This means creating 1 entry for every possible corner in the Hilbert Curve.


```r 
library(microbenchmark)
library(HilbertCurve)
library(tidyverse)
library(gghilbertstrings)
mb <- list()
for (i in 1:10) {
  df <- tibble(val = 1:4^i,
               size = runif(4^i, 1, 5),
               # create random sizes
               color = rep(c(1, 2, 3, 4), 4^(i - 1)))
  values <- df$val
  mb[[i]] <- microbenchmark(times = reps,
                     HilbertCurve = {
                       hc <- HilbertCurve(1, 4^i, level = i, newpage = FALSE)
                     },
                     gghilbertstrings = {
                       ggh <- hilbertd2xy(n = 4^i, values)
                     })
}
```



```{r setupmb, eval=TRUE, include=FALSE}

# Performance benchmark
library(microbenchmark)
library(HilbertCurve)
library(tidyverse)
library(gghilbertstrings)
mb <- list()
for (i in 1:10) {
  df <- tibble(val = 1:4^i,
               size = runif(4^i, 1, 5),
               # create random sizes
               color = rep(c(1, 2, 3, 4), 4^(i - 1)))
  values <- df$val
  mb[[i]] <- microbenchmark(times = reps,
                     HilbertCurve = {
                       hc <- HilbertCurve(1, 4^i, level = i, newpage = FALSE)
                     },
                     gghilbertstrings = {
                       ggh <- hilbertd2xy(n = 4^i, values)
                     })
}
res <- data.frame()
for (i in 1:length(mb)) {
  tmp <- mb[[i]] %>% as_tibble() %>% mutate(depth = i)  
  res <- res %>% bind_rows(tmp)
}

```

```{r output, echo=FALSE, message=FALSE, warning=FALSE}
library(scales)
res %>% mutate(time = time/1000) %>% 
  ggplot() + 
  aes(x = depth, y = time, color = expr) + 
  geom_jitter(width = 0.1, height = 0, alpha = 0.2) +
  geom_smooth() + 
  scale_y_log10(labels = comma) +
  scale_x_continuous(breaks = 3:14, minor_breaks = NULL) +
  scale_color_viridis_d(begin = 0.5, end = 0.8, option = "D") +
  labs(x = "Order of Hilbert Curve", y = "Time in ms (log-scale)", color = "Package",
       title = "Our package is two orders of magnitute faster",
       subtitle = glue("Comparison of {reps} repetitions across all orders."),
       caption = "Order 14 means 268,435,456 coordinates")
```



### Useful example

We use the `eliasdabbas/search-engine-results-flights-tickets-keywords` data set on [https://www.kaggle.com/eliasdabbas/search-engine-results-flights-tickets-keywords](Kaggle) as an example for a simple analysis. We map the full search URLs to the Hilbert curve and then add points when the URL was present for a specific search term. By comparing resulting facets we can see systematic difference in which provides show up for which search term.

```{r example_flt, echo=FALSE, message=FALSE, warning=FALSE}
#library(kaggler)
#kgl_auth() # requires a kaggle token.

data_set <- "eliasdabbas/search-engine-results-flights-tickets-keywords"
#refs <- kgl_datasets_list(owner_dataset = data_set)
#refs$datasetFiles
#kgl_datasets_download(owner_dataset = data_set, 
#                      fileName = refs$datasetFiles$name[1], datasetVersionNumber = 14)

read_plus <- function(flnm) {
    read_csv(flnm) %>% 
        mutate(filename = flnm)
}

df_with_sources <-
    list.files(path = "my_tests", 
               pattern = "*.csv", 
               full.names = T) %>% 
    map_df(~read_plus(.))


comparison <- df_with_sources$searchTerms %>% unique() %>% head(4)


df_with_sources %>%
  create_id_column(link) %>% 
  filter(searchTerms %in% comparison) %>% 
  gghilbertplot(gghid, 
                color = displayLink, 
                size = rank,
                label = displayLink,
                alpha = 0.1,
                jitter = 0.1,
                add_curve = F,
                curve_alpha = 0.1) +
  aes(group = displayLink) +
  facet_wrap(~searchTerms) +
  labs(title = "Comparsion of domains shown for different search queries") +
  guides(color = FALSE) -> p

p
#plotly::ggplotly(p)
```

Link: https://www.kaggle.com/eliasdabbas/search-engine-results-flights-tickets-keywords under License CC0

